Unrecognized States: Theory and Cases

### 2017-07-04 theory.tex
tightening through sec 2 and 3

We do not restrict the preferences and capabilities for players $g$ and $s$, with one exception. We assert that the payoffs for the party that cedes the issue of status (independence vs. reunification) are consistently low. This reflects the fact that the issue of status is indivisible and highly valued by each side and that many of the payments that could be offered are not credible (Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007; Schultz, 2010).

@@

\section{Explaining Outcomes of Secessionist Conflicts: The ``Status Quo'' Equilibrium} 
\label{sec:main}

Despite the preferences of the international community for peace, the most common outcome of secessionist conflicts in the post-WWII period has been reunification with the home state via outright military reconquest. By contrast, reunification through negotiated settlement has been very rare. In this section, we use a game theoretic model to explain the outcome that we find the most puzzling: perpetual unrecognized statehood.

@@

The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors. Player $c$ dislikes war and so will never invest in either state variable associated with winning since they increase the likelihood that one of the inside actors chooses to fight. It would also not invest in the government's payoffs from ceding. The patron will never invest in the government's payoffs from winning or the secessionists' payoffs from ceding, and it will not invest in the government's status quo payoffs because player $c$ will not invest in the government's payoffs from war so there is no need to counter such an investment. This leaves three state variables in which each outside actor might invest, which we address while defining the Status Quo Equilibrium.

@@

The strategies for the government and secessionists in this equilibrium are to play their best responses given the continuation values induced by the investments of the outside actors. Unless otherwise noted below, playing Status Quo is the best response for both inside actors in terms of continuation values.

The strategies for the outside actors in period $t$ are:

\begin{enumerate}
	\item The patron invests enough in the secessionists Status Quo payoffs to deter Player $c$ from investing in the secessionists' payoffs from Ceding. Otherwise, Player $c$ invests enough to induce the secessionists to play Cede.

	\item Player $c$ is prepared to counter the largest investment the patron is willing to make in the government's payoffs from ceding by making a counter-investment in the government's status quo payoffs. If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the government would play Cede.

	\item Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting. If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the secessionists would play Fight.
\end{enumerate}
\end{definition}

To understand the construction of this equilibrium, notice that there are three possible ways to disrupt the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order to establish that the Status Quo Equilibrium exists, it must be shown that each of these three deviations will be deterred.

@@

The international community's willingness to counteract investments by the patron toward the other two disturbances (i.e., Restriction (\ref{res:4})) implies that there will be no investments in equilibrium in cases (2) and (3). Case (3) also requires Restrictions (\ref{res:3}) of Definition 1 and the implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.

@@

First, note that the restrictions in Definition 1 are sufficient but not necessary. For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (\ref{res:6}) need not be met to maintain the status quo in the short run. The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.

On the other hand, the restrictions in Definition 1 also \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome. In fact, at least one additional equilibrium outcome always coexists with the status quo outcome. The equilibria may involve (Fight,Fight), (Fight,Cede) and/or (Cede,Fight) as outcomes of the stage game.

>>>>

We do not make any assumptions about the preferences and capabilities for players $g$ and $s$, with one exception. We assert that the payoffs for the party that cedes the issue of status (independence vs. reunification) are consistently low. This reflects the fact that the issue of status is indivisible and highly valued by each side and that many of the payments that could be offered are not credible (Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007; Schultz, 2010).

@@

\section{Explaining the ``Status Quo'' Equilibrium} 
\label{sec:main}

%Despite the preferences of the international community for peace, the most common outcome of secessionist conflicts in the post-WWII period has been reunification with the home state via outright military reconquest. By contrast, reunification through negotiated settlement has been very rare. In this section, we use a game theoretic model to explain the outcome that we find the most puzzling: perpetual unrecognized statehood.

@@

Of the six state variables, Player $c$ dislikes war and so will never invest in either state variable associated with winning since they increase the likelihood that one of the inside actors chooses to fight. It would also not invest in the government's payoffs from ceding. The patron will never invest in the government's payoffs from winning or the secessionists' payoffs from ceding, and it will not invest in the government's status quo payoffs because player $c$ will not invest in the government's payoffs from war so there is no need to counter such an investment. This leaves three state variables in which each outside actor might invest, which we address while defining the Status Quo Equilibrium.

@@

The strategies for the government and secessionists in this equilibrium are to play their best responses given the continuation values induced by the investments of the outside actors. Unless otherwise noted below, playing Status Quo is the best response for both inside actors.

The strategies for the outside actors in each period are:

\begin{enumerate}
	\item The patron invests enough in the secessionists Status Quo payoffs to deter Player $c$ from investing in the secessionists' payoffs from ceding. Otherwise, Player $c$ invests enough to induce the secessionists to play Cede.

	\item Potential investments by Player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding. If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the government would play Cede.

	\item Potential investments by Player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting. If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the secessionists would play Fight.
\end{enumerate}
\end{definition}

In order to establish that the Status Quo Equilibrium exists, we must show that each of three possible deviations will be deterred: (1) the secessionists provoked to Cede the international community, (2) the government provoked to cede by the patron, or (3) the secessionists provoked to fight by the patron.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.}

@@

The international community's willingness to counteract investments by the patron toward the other two disturbances (i.e., Restriction (\ref{res:4})) implies that there will be no investments in equilibrium in cases (2) and (3). Case (3) also requires Restriction (\ref{res:3}) of Definition 1 and the implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.

@@

The restrictions in Definition 1 do not provide for a unique equilibrium, or even a unique equilibrium outcome. At least one additional equilibrium outcome always coexists with the status quo outcome: at least one of (Fight,Fight), (Fight,Cede) and (Cede,Fight) will always be an equilibrium outcome of the stage game.
